An Extension of LTL with Rules\\ and its Application to Runtime Verification

### 2019-02-02 rule-ltl.tex
fixing problems

\noindent {\bf Related work.}
%
For propositional LTL, the logic ETL~\cite{WVS,Wolper} extends
the set of temporal operators 
by using automata (or, equivalently a right linear grammar). Each automaton defines a temporal operator,
where the input letters of the automaton correspond to future temporal subformulas. Each accepted run defines then the places
where the subformulas (corresponding to the occurrences of the input letters in the run) need to hold. Another expressive-equivalent extension, 
in~\cite{Wolper}, adds
state-dependent quantification over propositions to LTL.

@@

\klaus{You above introduce some concepts before they really are explained. I wonder whether the reader will understand. Specifically the E, but also the Q. Why not just introduce as we go along?}

@@

@@

\noindent {\bf Evaluation order.} The order of calculating $\current$ is subtle here, because
auxiliary variables can appear both on the left and right of their
definition. For PLTL, one can start calculating the subformulas
bottom up or top down. But for EPLTL, this would not work. To see this, consider, for example, the formula~(\ref{form3}). It contains
the definition $q := \ominus \neg q$. Now, we cannot calculate this
bottom up, since the left-hand $q$ is not calculated yet.
If we calculate it top down, we will encounter the same problem.
However, notice that the top down calculation does not need the
value of $q$ to calculate $\ominus \neg q$; this value should already be available as $\old ( \neg q )$.

@@

\noindent {\bf Syntax.} 
The formulas of the core FLTL logic are 
defined by the following grammar,
where $p$ denotes to a relation $R$,
$a$ is a constant representing a value in $d ( R )$ and $x$ is a variable with $\domain ( x ) = d ( p )$.
\begin{center}
$\varphi ::= \true  \; | \;
    p ( a ) \; | \;
    p ( x ) \; | \;
    ( \varphi \wedge \varphi ) \;  |   \;
   \neg \varphi \; | \;
   \bigcirc \varphi \; | \; 
   ( \varphi \; \until \; \varphi ) \; | $ \\
   $ \ominus \; \varphi \; | \;
    ( \varphi  \; \since  \; \varphi ) \; | \;
    \exists x \; \varphi$
\end{center}

@@

\noindent {\bf Semantics.} 
A model is a set of
 relations ${\cal R} = \{ R_1 \, \ldots , R_m \}$, where each of these relations has a last parameter $i$ that is a natural number.

It it standard, and more intuitive (in particular for runtime verification) to cast models of temporal logic in terms of sequences states rather than temporal relations. To do that,
 we denote by $R_k [ i ]$ the restriction of
 the relation $R_k$ to the value $i$ of the last parameter. A model ${\cal R}$ is then a sequence ${\cal R} [ 1 ]\, {\cal R} [ 2 ] \ldots$, where
 each {\em state} ${\cal R} [ i ] = {\cal R} [i] = \{ R_1 [i], \ldots , R_m [ i ] \}$  (a set of relations).
 The denotation $p_k$ will correspond to the relation $R_k$.

>>>>

\subsubsection{Related Work}

For propositional LTL, the logic ETL~\cite{WVS,Wolper} extends
the set of temporal operators 
by using automata (or, equivalently a right linear grammar). Each automaton defines a temporal operator,
where the input letters of the automaton correspond to future temporal subformulas. Each accepted run defines then the places
where the subformulas (corresponding to the occurrences of the input letters in the run) need to hold. Another expressive-equivalent extension, 
in~\cite{Wolper}, adds
state-dependent quantification over propositions to LTL.

@@

\klaus{You above introduce some concepts before they really are explained. Q and E. I understand why you do this (to have it in one place), but it may be too difficult for the reader to grasp if not "initiated".
Especially the E.}

\klaus{We use ELTL (etc) to denote Extended LTL. 
But QLTL is also an extension of LTL. These logics 
are all extensions of LTL. Perhaps RLTL (for Rule-based 
LTL) would have been better ... or something. Not urgent, 
we can always change that if we want to.}

@@

\subsubsection{Quantified LTL}

@@

\subsubsection{Evaluation Order}

The order of calculating $\current$ is subtle here, because
auxiliary variables can appear both on the left and right of their
definition. For PLTL, one can start calculating the subformulas
bottom up or top down. But for EPLTL, this would not work. To see this, consider, for example, the formula~(\ref{form3}). It contains
the definition $q := \ominus \neg q$. Now, we cannot calculate this
bottom up, since the left-hand $q$ is not calculated yet.
If we calculate it top down, we will encounter the same problem.
However, notice that the top down calculation does not need the
value of $q$ to calculate $\ominus \neg q$; this value should already be available as $\old ( \neg q )$.

@@

\subsubsection{Syntax}

The formulas of the core FLTL logic are 
defined by the following grammar,
where $p$ denotes to a relation $R$,
$a$ is a constant representing a value in $d ( R )$ and $x$ is a variable with $\domain ( x ) = d ( p )$.
\begin{center}
$\varphi ::= \true  \; | \;
    p ( a ) \; | \;
    p ( x ) \; | \;
    ( \varphi \wedge \varphi ) \;  |   \;
   \neg \varphi \; | \;
   \bigcirc \varphi \; | \; 
   ( \varphi \; \until \; \varphi ) \; | $ \\
   $ \ominus \; \varphi \; | \;
    ( \varphi  \; \since  \; \varphi ) \; | \;
    \exists x \; \varphi$
\end{center}

@@

\subsubsection{Semantics}

A model is a set of
 relations ${\cal R} = \{ R_1 \, \ldots , R_m \}$, where each of these relations has a last parameter $i$ that is a natural number. 
% 
It it standard, and more intuitive (in particular for runtime verification) to cast models of temporal logic in terms of sequences states rather than temporal relations. To do that,
 we denote by $R_k [ i ]$ the restriction of
 the relation $R_k$ to the value $i$ of the last parameter. A model ${\cal R}$ is then a sequence ${\cal R} [ 1 ]\, {\cal R} [ 2 ] \ldots$, where
 each {\em state} ${\cal R} [ i ] = {\cal R} [i] = \{ R_1 [i], \ldots , R_m [ i ] \}$  (a set of relations).
 The denotation $p_k$ will correspond to the relation $R_k$.
